---
title: "Area of habitat from hand-drawn map"
author: "barrett wolfe"
output:
  html_document:
    df_print: paged
---

The goal is to automatically detect hand-drawn regions on a hand drawn map and
calculate the area. 

Area will be calculated in pixels, to convert to pixels I'd recommend manually
measuring a feature of known dimensions in pixels (doable in R but you'd 
probably be better off using imageJ), and then it is a 
simple conversion from pixels to whatever units of area.  

Also, this works if the map can be considered planar (i.e, it explicitly is
or if not the extent is small enough the error will be negligible)




```{r include = FALSE}
library(imager)
library(scales)
library(Momocs)

```
####First load this beautiful hand drawn map:


```{r fig.width = 10, fig.height = 10}
img <- load.image("areas.jpg")
plot(img)
```

####Convert to grayscale for processing:  

```{r}
img.g <- grayscale(img)
plot(img.g)
```

#### Thresholding to detect polygons:  

####You want the contrast of ink vs paper as high as possible when scanning. Also, you likely
will need to tweak the  `adjust` parameter to get the best possible seperation.
If you still aren't getting good separation of images and background, you might
need to do pre-processing in other software.  

`adjust` parameter too high:  

```{r}
plot(threshold(img.g, adjust = 1))
```

`adjust` too low: 
```{r}
plot(threshold(img.g, adjust = 0.6))
```

`adjust` just right: 
```{r}
plot(threshold(img.g, adjust = 0.89))
```
###Some artifact like in the lower left isn't an issue if it isn't interfering with
the areas of interest on the map, better that than not having solid outlines.  

Now to separate the individual areas:

```{r}
area.list <- split_connected(threshold(img.g, adjust = 0.89))
```

####Plotting the whole list will show you if you've captured your areas successfully,
the main thing will be that any artifact are smaller than the smallest areas of interest.
That way you can filter by the size of the areas (number of pixels) to retain only the 'real' shapes.  

```{r, fig.width = 12, fig.height = 12}
plot(area.list, xlim = c(100, 1100), ylim = c(500,1400))
sapply(area.list, \(x)sum(x))
```

####Looks like 4000 is our cutoff for a real shape, there are quite a few smaller
ones detected but they are just artifacts. If everything is working, we should
recognise these shapes we have retained:  
```{r, fig.width = 10, fig.height = 10}
real.areas <- area.list[which(sapply(area.list, function(x) sum(x)>4000))]
#in case the main area is not the first, sort by descending order of size
real.areas <- real.areas[order(sapply(real.areas, sum), decreasing = TRUE)]
plot(real.areas,xlim = c(0, 1200), ylim = c(400,1500))
```

####Finally, we want the area of each, and for verification we want to print the
area in an outline of each shape. To extact outlines, we need to convert the
shapes to masks (black pixels on empty background) and identify an xy point
inside each shape.  

```{r fig.height = 10, fig.width = 8}

#centroid of each shape, if it just so happens there is a smaller shape at
#the middle of the main map areas, may just need to provide a coordinate manually
#also if the outlines later on are flipped, switch the commented line in the
#function below

centroids <-
  round(coo_centpos(Out(lapply(real.areas, function(x) {
    which(apply(as.matrix(x),1,rev), arr.ind = TRUE)
#   which(as.matrix(x), arr.ind = TRUE)
      }))),
  0)

#also for whatever reason these were plotting in mirrored locations, just for 
#simplicity I'll just flip them for plotting
label_pos <-
  round(coo_centpos(Out(lapply(real.areas, function(x) {
    which(apply(apply(as.matrix(x),1,rev),2,rev), arr.ind = TRUE)
#   which(as.matrix(x), arr.ind = TRUE)
      }))),
  0)

#inverting shapes
real.inv <- as.imlist(lapply(real.areas, \(x)!x))


outlines <- mapply(\(x,y) {
  x1 <- as.matrix(as.cimg(x))
 #if outlines are flipped incorrectly, remove the following line
   x1 <- apply(x1, 1, rev)
import_Conte(x1,centroids[y,])
    },
  x = real.inv,  y= 1:nrow(centroids))




plot(img, xlim = c(0,1200), ylim = c(1600,0))
sapply(outlines, \(x) {
  lines(x[,1],x[,2],col = alpha("red3", 0.5), lwd = 2)
  })

text(label_pos[,2], label_pos[,1],
     labels = sapply(real.areas, function(x) sum(x)),
     col = "red4")
```

####That is a bit busy so maybe we are better off just numbering them so they can
be matched to the area in a vector. Also, I'll remove the area of the 'background'
and instead include the total area (sum of shapes and background).  



```{r fig.height = 10, fig.width= 8 }
plot(img, xlim = c(0,1200), ylim = c(0,1600))
sapply(outlines, \(x) {
  lines(x[,1],x[,2],col = alpha("red3", 0.5), lwd = 2)
  })

text(label_pos[-1,2], label_pos[-1,1],
     labels = 2:length(real.areas),
     col = "red4", cex = 2)

area_tab <- data.frame(Region = 1:length(real.areas),
                       Area_in_pixels = sapply(real.areas, function(x) sum(x))
)
area_tab[nrow(area_tab)+1,] <- c("Total", sum(area_tab$Area_in_pixels))

knitr::kable(area_tab[-1,], col.names = gsub("_", " ", names(area_tab)))
```
